Efficient Machine Learning Approaches for Fast Risk Evaluation of VAs

### 2025-01-07 pre_private/pre.tex
new

% use this so appendices' page numbers do not count
\usepackage{appendixnumberbeamer}
\usepackage{booktabs}
\usepackage{subfigure}
\usepackage[round]{natbib}
\usepackage{mathtools}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{bm}
\usepackage{multirow}
\usepackage{graphicx}

@@

\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\I}{\mathbbm{1}}

\newcommand{\fracM}{\frac{1}{M}}
\newcommand{\fracMM}{\frac{1}{M^2}}
\newcommand{\sumM}{\sum_{i=1}^M}

\newcommand{\fracN}{\frac{1}{N}}
\newcommand{\fracNN}{\frac{1}{N^2}}
\newcommand{\sumN}{\sum_{j=1}^N}

\newcommand{\rhoh}{\hat{\rho}}
\newcommand{\Lh}{\hat{L}}
\newcommand{\betah}{\hat{\beta}}

\newcommand{\Bias}{\textnormal{Bias}}
\newcommand{\Variance}{\textnormal{Variance}}

\newcommand{\SNS}{\textnormal{SNS}}
\newcommand{\REG}{\textnormal{REG}}
\newcommand{\KS}{\textnormal{KS}}
\newcommand{\LR}{\textnormal{LR}}
\newcommand{\KRR}{\textnormal{KRR}}
\newcommand{\skipv}{\vspace{8pt}}

@@

    \begin{figure}[H]
        \makebox[\textwidth][c]{
            \subfigure{
                \begin{tikzpicture}[scale=0.7]
                \draw[-latex] (0,0)--(2,0);
                \draw[-latex] (2.2,0)--(4,2) node[right]  {$\blacksquare$};
                \draw[-latex] (2.2,0)--(4,0.5) node[right]  {$\blacksquare$};
                \draw[-latex] (2.2,0)--(4,-2) node[right]  {$\blacksquare$};
                \node at (1.2,-0.3) {$S_i$};
                \node at (3.8,-0.7) {$\vdots$};
                \node at (4.2,-0.7) {$\vdots$};
                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (4.5,2.2) --  (4.5,-2.2);
                \node at (5.3,0) {$\hat L_i$};
                \node at (2,-3.1) {1 Outer Path for Options};
                \end{tikzpicture}}
            \subfigure{
                \begin{tikzpicture}[scale=0.6]
                \draw[-latex] (0,0)--(10,0);
                \draw[-latex] (0,0)--(-1,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (0,0)--(0,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (0,0)--(1,1.5) node[above] {$\blacksquare$};
                \node at (0.5,1.2) {\small$\dots$};
                \node at (0.5,1.85) {\small$\dots$};

                \draw[-latex] (3,0)--(2,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (3,0)--(3,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (3,0)--(4,1.5) node[above] {$\blacksquare$};
                \node at (3.5,1.2) {\small$\dots$};
                \node at (3.5,1.85) {\small$\dots$};

                \draw[-latex] (6,0)--(5,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (6,0)--(6,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (6,0)--(7,1.5) node[above] {$\blacksquare$};
                \node at (6.5,1.2) {\small$\dots$};
                \node at (6.5,1.85) {\small$\dots$};

                \draw[-latex] (9,0)--(8,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (9,0)--(9,1.5) node[above] {$\blacksquare$};
                \draw[-latex] (9,0)--(10,1.5) node[above] {$\blacksquare$};
                \node at (9.5,1.2) {\small$\dots$};
                \node at (9.5,1.85) {\small$\dots$};

                \node at (7.5,0.7) {\Huge $\dots$};

                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (-1,2) --  (1,2);
                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (2,2) --  (4,2);
                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (5,2) --  (7,2);
                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (8,2) --  (10,2);
                \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=12pt}] (0,2.4) --  (9,2.4);
                \node at (4.5,-0.5) {$S_i$};
                \node at (4.5,3.5) {$\hat L_i$};
                \node at (4.5,-1.3) {1 Outer Path for VAs};
                \end{tikzpicture}
                }}
	\end{figure}

@@

\textbf{Key advantages:}
\begin{itemize}
    \item more efficient than a two-stage procedure;
    \item avoids specifying $m$;
\end{itemize}

@@

    Consider estimating the $95\%$ CVaR with 100,000 outer scenarios.
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=1]
            % Make the height of the solid rectangle half.
            \draw[line width=1pt, fill=gray, fill opacity=0.40] (0,0) rectangle (3,1); 
            % Adjust node positions for the new height.
            \node[above] at (1.5, 0.5) {5\% \textbf{true} tails}; 
            \node[above] at (1.5, 0) {(5,000 scenarios)};
            % Adjust the dashed lines for the new height.
            \draw[dashed, line width=1pt] (3,0) -- (10,0); 
            \node[above] at (6.5, 0.3) {non-tails};
            \draw[dashed, line width=1pt] (3,1) -- (10,1); 
            \node[above] at (9.5, 0.3) {...};
        \end{tikzpicture}
    \end{figure}
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=1]
            % Make the height of the solid rectangle half.
            \draw[line width=1pt, fill=gray, fill opacity=0.20] (0,0) rectangle (3,1);
            % Adjust node positions for the new height.
            \node[above] at (1.5, 0.5) {5\% \textbf{predicted} tails};
            \node[above] at (1.5, 0) {(5,000 scenarios)};
            % Adjust the dashed boxes for the new height.
            \draw[dashed, line width=1pt, fill=gray, fill opacity=0.10] (3,0) rectangle (5,1);
            \node[above] at (4, 0.5) {+2\% margin};
            \node[above] at (4, 0) {(2,000 more)};
            \draw[dashed, line width=1pt, fill=gray, fill opacity=0.10] (5,0) rectangle (7,1);
            \node[above] at (6, 0.5) {+2\% margin};
            \node[above] at (6, 0) {(2,000 more)};
            % Adjust the dashed lines for the new height.
            \draw[dashed, line width=1pt] (7,0) -- (10,0);
            \draw[dashed, line width=1pt] (7,1) -- (10,1);
            \node[above] at (9.5, 0.3) {...};
        \end{tikzpicture}
    \end{figure}

@@

\section{Transfer Learning for Dynamic Hedging}
\begin{frame}{Transfer Learning for Dynamic Hedging}

>>>>

% use this so appendices' page numbers do not count
\usepackage{appendixnumberbeamer}
\usepackage{booktabs}
\usepackage[round]{natbib}
\usepackage{mathtools}
% \usepackage{subfigure}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{bm}

@@

@@

    \begin{figure}[H]
            \includegraphics[width=0.33\textwidth]{./tikz/illustration_spns.pdf}
            \includegraphics[width=0.62\textwidth]{./tikz/illustration_mpns.pdf}
	\end{figure}

@@

\textbf{Key advantages:}
\begin{itemize}
    \item more efficient than a two-stage procedure;
    \item avoids specifying $m$.
\end{itemize}

@@

    Consider estimating the $95\%$ CVaR with 100,000 outer scenarios.
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=1]
            % Make the height of the solid rectangle half.
            \draw[line width=1pt, fill=gray, fill opacity=0.40] (0,0) rectangle (3,1); 
            % Adjust node positions for the new height.
            \node[above] at (1.5, 0.5) {5\% \textbf{true} tails}; 
            \node[above] at (1.5, 0) {(5,000 scenarios)};
            % Adjust the dashed lines for the new height.
            \draw[dashed, line width=1pt] (3,0) -- (10,0); 
            \node[above] at (6.5, 0.3) {non-tails};
            \draw[dashed, line width=1pt] (3,1) -- (10,1); 
            \node[above] at (9.5, 0.3) {...};
        \end{tikzpicture}
    \end{figure}
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=1]
            % Make the height of the solid rectangle half.
            \draw[line width=1pt, fill=gray, fill opacity=0.20] (0,0) rectangle (3,1);
            % Adjust node positions for the new height.
            \node[above] at (1.5, 0.5) {5\% \textbf{predicted} tails};
            \node[above] at (1.5, 0) {(5,000 scenarios)};
            % Adjust the dashed boxes for the new height.
            \draw[dashed, line width=1pt, fill=gray, fill opacity=0.10] (3,0) rectangle (5,1);
            \node[above] at (4, 0.5) {+2\% margin};
            \node[above] at (4, 0) {(2,000 more)};
            \draw[dashed, line width=1pt, fill=gray, fill opacity=0.10] (5,0) rectangle (7,1);
            \node[above] at (6, 0.5) {+2\% margin};
            \node[above] at (6, 0) {(2,000 more)};
            % Adjust the dashed lines for the new height.
            \draw[dashed, line width=1pt] (7,0) -- (10,0);
            \draw[dashed, line width=1pt] (7,1) -- (10,1);
            \node[above] at (9.5, 0.3) {...};
        \end{tikzpicture}
        \caption{Illustration: a safety margin of $4\%$ ($m = 9000$)}
    \end{figure}

@@

\section{Transfer Learning for Rapid Adaptation of DNN Metamodels}
\begin{frame}{Transfer Learning for Rapid Adaptation of DNN Metamodels}

    \textbf{Challenge:} Adapting deep neural network metamodels to changing conditions.
    \begin{itemize}
        \item   Retraining from scratch is computationally expensive.
        \item   Efficient incorporation of new VA contract data.
        \item   Balancing model accuracy and computational costs.
    \end{itemize}

    \textbf{Solution:} Transfer learning (TL) to develop adaptable, efficient metamodels for VA dynamic hedging.

    \begin{itemize}
        \item   Pre-train deep neural network on contracts with abundant simulation data.
        \item   Fine-tune on smaller dataset of new contracts/market conditions.
        \item   Leverages shared features between VA contracts.
        \item   Computational savings:
            \begin{itemize}
                \item   Reduced training time.
                \item   Fewer data points needed for good performance.
            \end{itemize}
    \end{itemize}

\begin{frame}{Transfer Learning Framework}

    \textbf{Key Components:}
    \begin{itemize}
        \item   \textbf{Domain} $\mathcal{D}$: feature space $\mathcal{X}$ + probability distribution $F$
        \item   \textbf{Task} $\mathcal{T}$: label space $\mathcal{Y}$ + predictive function $f: \mathcal{X} \rightarrow \mathcal{Y}$
    \end{itemize}

    \vspace{10pt}
    \textbf{Source vs. Target:}
    \begin{itemize}
        \item   \textbf{Source:} $\mathcal{D}_{\text{So}} = \{\mathcal{X}_{\text{So}}, F_{\text{So}}(X)\}$
        \item   \textbf{Target:} $\mathcal{D}_{\text{Ta}} = \{\mathcal{X}_{\text{Ta}}, F_{\text{Ta}}(X)\}$
    \end{itemize}

    \vspace{10pt}
    \textbf{Our Goal:}
    \begin{itemize}
        \item   \textbf{Input features} $X$: risk factors from outer simulation
        \item   \textbf{Output labels} $L$: contract losses
        \item   \textbf{Source and target}: from VAs with abundant simulation data to new VAs with limited data
        \item   \textbf{Goal}: improve $f_{\text{Ta}}(\cdot)$ using knowledge from $\mathcal{D}_{\text{So}}$ and $f_{\text{So}}(\cdot)$
    \end{itemize}

\begin{frame}{Transfer Learning Techniques}

    \textbf{Common Techniques:}
    \begin{itemize}
        \item   \textbf{Fine-tuning:} a model pre-trained on a source task is used as a starting point for a target task.
        \item   \textbf{Layer freezing:} only part of the model is fine-tuned.
        \item   \textbf{Multi-task learning:} perform training on multiple tasks simultaneously.
    \end{itemize}

    \textbf{Key considerations:}
    \begin{itemize}
        \item   Similarity between source and target tasks
        \item   Appropriate learning rate
    \end{itemize}

\begin{frame}{Fine-tuning Algorithm for LSTM Metamodels in VA Hedging}

    \begin{algorithm}[H]
        \caption{Fine-tuning Algorithm for LSTM Metamodels in VA Hedging}
        \begin{algorithmic}[1]
            \STATE \textbf{Input:} $\mathcal{D}_{\text{So}} = \{(X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)})\}_{i=1}^{M_{\text{So}}}$ , $\mathcal{D}_{\text{Ta}} = \{(X_{\text{Ta}}^{(i)}, L_{\text{Ta}}^{(i)})\}_{i=1}^{M_{\text{Ta}}}$, $\alpha_{\text{So}}$, and $\alpha_{\text{Ta}}$.
            \STATE Train a LSTM metamodel $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$:
            \begin{equation*}
                \theta_{\text{So}} = \min_{\theta} \frac{1}{M_{\text{So}}} \sum_{i=1}^{M_{\text{So}}} \left( f_{\text{So}}(X_{\text{So}}^{(i)}; \theta) - L_{\text{So}}^{(i)} \right)^2
            \end{equation*}
            \STATE Initialize the target metamodel parameters $\theta_{\text{Ta}}$ using the pre-trained metamodel parameters:
            \begin{equation*}
                \theta_{\text{Ta}} \gets \theta_{\text{So}}
            \end{equation*}
            \STATE Fine-tune the entire LSTM metamodel $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ on the target dataset $\mathcal{D}_{\text{Ta}}$ using a smaller learning rate $\alpha_{\text{Ta}}$:
            \begin{equation*}
                \theta_{\text{Ta}} = \min_{\theta} \frac{1}{M_{\text{Ta}}} \sum_{i=1}^{M_{\text{Ta}}} \left( f_{\text{Ta}}(X_{\text{Ta}}^{(i)}; \theta) - L_{\text{Ta}}^{(i)} \right)^2
            \end{equation*}
            \STATE \textbf{Output:} Final adapted LSTM metamodel $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ for the target task
        \end{algorithmic}
    \end{algorithm}

\begin{frame}{Layer Freezing Algorithm for LSTM Metamodels in VA Hedging}

    \begin{algorithm}[H]
        \caption{Layer Freezing Algorithm for LSTM Metamodels in VA Hedging}
        \begin{algorithmic}[1]
            \STATE \textbf{Input:} $\mathcal{D}_{\text{So}} = \{(X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)})\}_{i=1}^{M_{\text{So}}}$ , $\mathcal{D}_{\text{Ta}} = \{(X_{\text{Ta}}^{(i)}, L_{\text{Ta}}^{(i)})\}_{i=1}^{M_{\text{Ta}}}$, $\alpha_{\text{So}}$, and $\alpha_{\text{Ta}}$.
            \STATE Train LSTM model $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$.
            \STATE Initialize the target model parameters $\theta_{\text{Ta}} = [\theta_0, \theta_1]$ using the pre-trained source model parameters $\theta_{\text{So}}$:
            \begin{equation*}
                \theta_{\text{Ta}} \gets \theta_{\text{So}} = [\theta_0, \theta_1]
            \end{equation*}
            \STATE Freeze the parameters of the shared layers $\theta_0$ and fine-tune the trainable layers $\theta_1$ on the target dataset $\mathcal{D}_{\text{Ta}}$:
            \begin{equation*}
                \theta_{\text{Ta}} = \min_{\theta_1} \frac{1}{M_{\text{Ta}}} \sum_{i=1}^{M_{\text{Ta}}} \left( f_{\text{Ta}}(X_{\text{Ta}}^{(i)}; [\theta_0, \theta_1]) - L_{\text{Ta}}^{(i)} \right)^2
            \end{equation*}
            \STATE \textbf{Output:} Adapted model $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ for the target task.
        \end{algorithmic}
    \end{algorithm}

\begin{frame}{Multi-task Learning Algorithm for LSTM Metamodels in VA Hedging}

    \begin{algorithm}[H]
        \caption{Multi-task Learning Algorithm for LSTM Metamodels in VA Hedging}
        \begin{algorithmic}[1] \label{alg3:multiTaskLearning}
            \STATE \textbf{Input:} learning rate $\alpha$, set of $K$ tasks $\{\mathcal{T}_k\}_{k=1}^K$ with datasets $\mathcal{D}_k = \{(X_k^{(i)}, L_k^{(i)})\}_{i=1}^{M_k}$, task-specific parameters $\theta_k$ for each task $k$, and shared parameters $\theta_0$.

            \STATE Train the multi-head LSTM metamodel on all $K$ tasks simultaneously by minimizing the multi-task loss function:
            \begin{equation} \label{eq3:multiTaskLoss}
                \min_{\theta_0, \{\theta_k\}_{k=1}^K} \sum_{k=1}^K \frac{1}{M_k} \sum_{i=1}^{M_k} \left( f_i(X_k^{(i)}; \theta_0, \theta_k) - L_k^{(i)} \right)^2
            \end{equation}

            \STATE Update both the shared parameters $\theta_0$ and task-specific parameters $\{\theta_k\}_{k=1}^K$ simultaneously using backpropagation and gradient descent with learning rate $\alpha$.

            \STATE \textbf{Output:} Trained multi-task metamodel $f(\cdot; \theta_0, \{\theta_k\}_{k=1}^K)$ for all $K$ tasks
        \end{algorithmic}
    \end{algorithm}

\begin{frame}{Experiment Setup}

    \begin{table}[ht!] 
        \centering
        \begin{tabular}{lcccc} 
        \toprule
        \textbf{Contract} & \textbf{Asset Model} & \textbf{Lapse} & \textbf{$M_{\text{So}}$}  & \textbf{$M_{\text{Ta}}$}\\
        \midrule
        GMMB & GBM & No lapse & 50000 & N/A \\
        GMMB & RS-GBM & No lapse & 50000 & 2000 \\
        GMMB & RS-GBM & Static lapse & 50000 & 2000 \\
        GMMB & RS-GBM & Dynamic lapse & 50000 & 2000 \\
        GMWB & RS-GBM & Dynamic lapse & N/A & 2000 \\
        \bottomrule
        \end{tabular}
        \caption{VA Contracts for Transfer Learning Experiments}
    \end{table}

    We aim to examine the performance of TL techniques
    \begin{itemize}
        \item   learning the lapse features,
        \item   learning the dynamic lapse, and 
        \item   transferring to other contract types.
    \end{itemize}

\begin{frame}{Learning Lapse Features}
    \begin{figure}[ht!]
        \centering
        \begin{subfigure}{0.48\textwidth}
            \includegraphics[width=\textwidth]{../project3/figures/figure1a.png}
            \caption{Extensive Training on Target Task}
        \end{subfigure}
        \begin{subfigure}{0.48\textwidth}
            \includegraphics[width=\textwidth]{../project3/figures/figure1b.png}
            \caption{Without TL}
        \end{subfigure}
        \caption{Direct training on RS-GBM GMMB with static lapse}
    \end{figure}
\end{frame}

\begin{frame}{Learning Lapse Features}
    \begin{figure}[ht!]
        \centering
        \begin{subfigure}{0.48\textwidth}
            \includegraphics[width=\textwidth]{../project3/figures/figure1c.png}
            \caption{With Fine-tuning}
        \end{subfigure}
        \begin{subfigure}{0.48\textwidth}
            \includegraphics[width=\textwidth]{../project3/figures/figure1d.png}
            \caption{With Layer Freezing}
        \end{subfigure}
        \caption{TL on RS-GBM GMMB with static lapse}
    \end{figure}
